---
title: Identifying key species habitats and corridors
author: Lukmann Haqeem
date: "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: scholar.bib
header-includes:
    \usepackage{float}
output:
     pdf_document: default
     html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
\
\

## Getting started

```{r packages, warning = FALSE, message = FALSE}
# Load packages
library(unmarked)
library(jagsUI)
library(raster)
library(gdistance)
library(cluster)
library(factoextra)

# Load data sets
#load sample data
set.seed(1234)
dat <- read.csv(file.choose(), comment = "#")
head(dat)
dim(dat)
#load study area (for predictions part later)
data(Switzerland)
head(Switzerland)
dim(Switzerland)

# Plot data sets
#study area
raster <- rasterFromXYZ(data.frame(x = Switzerland$x,
								   y = Switzerland$y,
								   z = Switzerland$elevation))
mapPalette <- colorRampPalette(c("grey88", "grey66", "grey44", "grey22"))
plot(raster, col = mapPalette(100), main = "Location of sampling points")
#sampling points
points(dat$easting, dat$northing, pch = 0, cex = 1, col = "red") 

# Perform EDA
#extract the count data
C <- as.matrix(dat[,3:5])
#check initial patterns or anomalies
head(C)
Cmax <- apply(C, 1, max, na.rm = TRUE)
min(Cmax)
max(Cmax)
nSites <- nrow(C)
nOcc <- rowSums(!is.na(C))
```
\

## Statistical modeling

```{r jags, warning = FALSE, message = FALSE}
# Run Bayesian N-mixture model 
#construct JAGS building block for state-space model
modelText <- "
model{
for(i in 1:nSites){
    # Likelihood
    log(lambda[i]) <- b0 +
                      bForest * forest[i] +
                      bForest2 * forest2[i] +
                      bElevation * elevation[i] +
                      bElevation2 * elevation2[i]                     
    # Ecological process
    N[i] ~ dpois(lambda[i]) 
    
    for (j in 1:nOcc[i]){
        # Observation process
        C[i,j] ~ dbin(p, N[i]) 
    }
}
    # Priors
    b0 ~ dunif(-10, 10)          
    bForest ~ dunif(-5, 5)       
    bForest2 ~ dunif(-5, 5)      
    bElevation ~ dunif(-5, 5)    
    bElevation2 ~ dunif(-5, 5)   
    p ~ dunif(0, 1)
}
 "
writeLines(modelText, "model1.jags")
#organise data sets for the model 
jagsData <- list(nSites = nrow(C),
                 nOcc = nOcc,
                 C = C,
                 forest = dat$forest,
                 forest2 = dat$forest^2,
                 elevation = dat$elevation,
                 elevation2 = dat$elevation^2)
str(jagsData)             
#set parameters to monitor 
monitor <- c("b0", "bForest", "bForest2", "bElevation", "bElevation2")
#set initial values for parameters
init <- function() list(N = Cmax, z = rep(1, nrow(C)))

#mcmc settings
nc <- 3
ni <- 30000
nb <- ni/2
na <- 1000

#run the model
system.time(
    sim1 <- jags(data = jagsData,
                 inits = init,
                 parameters.to.save = monitor,
                 model.file = "model1.jags", 
                 n.chains = nc,
                 n.iter = ni,
                 n.burn = nb, 
                 n.adapt = na,
                 parallel = TRUE,
                 DIC = FALSE)
) # 22 s
#check for chains convergence               
mean(unlist(sim1$Rhat) > 1.1)
```
\

## Making prediction

```{r predict, eval = FALSE, warning = FALSE, message = FALSE}
# Using the model for making predictions across the entire study area (~SDM)
#estimate 1000 abundance estimates for the entire study area  
tmp <- matrix(NA, nrow = n, ncol = nrow(Switzerland))
n <- 1000
ord <- sample(1:length(sim1$sims.list$b0), n, replace = FALSE)
system.time(
	for(i in 1:n){
		for(j in 1:nrow(Switzerland)){
		tmp[i,j] <- exp(sim1$sims.list$b0[ord[i]] +
						sim1$sims.list$bForest[ord[i]]*Switzerland$forest[j] +
						sim1$sims.list$bForest2[ord[i]]*Switzerland$forest[j]^2 +
						sim1$sims.list$bElevation[ord[i]]*Switzerland$elevation[j] +
						sim1$sims.list$bElevation2[ord[i]]*Switzerland$elevation[j]^2)
		}
	}
) # 256 sec
#save(tmp, file = "1000AbundanceLayers.RData")
```

```{r kmeans, warning = FALSE, message = FALSE}
# Estimate optimal number of clusters 
#load prediction
load(file = "1000AbundanceLayers.RData")
dim(tmp)

#compare elevation and species abundance maps
Switzerland$abundance <- colMeans(tmp)
raster_abundance <- rasterFromXYZ(data.frame(x = Switzerland$x, y = Switzerland$y, z = Switzerland$abundance))
par(mfrow = c(1, 2))
plot(raster, col = mapPalette(100), main = "Elevation map")
plot(raster_abundance, main = "Abundance map")
par(mfrow = c(1, 1))

#'identify' key areas  
high_abundance <- Switzerland[Switzerland$abundance > 4.5,] # say, greater than or equal to this is considered high species abundance
raster <- rasterFromXYZ(data.frame(x = Switzerland$x, y = Switzerland$y, z = Switzerland$elevation))
plot(raster, col = mapPalette(100), main = "High abundance areas")
points(high_abundance$x, high_abundance$y, pch = 16, cex = 1, col = adjustcolor("orange", 0.1)) 

#calculate optimal number of clusters and centroids 
#calculate k-means clusters
system.time(
	gap_stat <- clusGap(high_abundance, FUN = kmeans, nstart = 25, K.max = 10, B = 100, iter.max = 50)
) # 85 s
fviz_gap_stat(gap_stat)
#calculate centroids
A <- matrix(NA, nrow = length(high_abundance$x), ncol = 2)
A[,1] <- high_abundance$x
A[,2] <- high_abundance$y
distance <- 119500	# adjust until n. centroids = optimal n. clusters
chc <- hclust(dist(data.frame(x = high_abundance$x, y = high_abundance$y)), method = "complete")
chcd <- cutree(chc, h = distance) 
centroids <- aggregate(A, list(cluster = chcd), mean)
pts <- cbind(x = centroids[,2], y = centroids[,3])
plot(raster, col = mapPalette(100), main = "Location of centroids")
points(high_abundance$x, high_abundance$y, pch = 16, cex = 1, col = adjustcolor("orange", 0.1)) 
text(pts[,1], pts[,2], label = LETTERS[1:nrow(pts)], col = "green")
```

```{r analysis1, warning = FALSE, message = FALSE}
# Aligning corridors on the map
#align corridors with elevation information only
plot(raster, col = mapPalette(100), main = "Corridors without wildlife information")
raster2 <- rasterFromXYZ(data.frame(x = Switzerland$x,
								   y = Switzerland$y,
								   z = Switzerland$elevation))
f <- function(x) max(x) - x[1] + x[2] 
tr <- transition(raster2, transitionFunction = f, directions = 8)
slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
points(high_abundance$x, high_abundance$y, pch = 16, cex = 1, col = adjustcolor("orange", 0.1)) 
for(j in 1:nrow(centroids)){
	for(k in 1:nrow(centroids)){
		plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
	  	     col = adjustcolor("cyan", 1), lwd = 1, add = TRUE)
	}
}
text(pts[,1], pts[,2], label = LETTERS[1:nrow(pts)], col = "green")

#align corridors with abundance estimate (only posterior mean)
plot(raster, col = mapPalette(100), main = "Corridors with wildlife information (post. mean only)")
raster3 <- rasterFromXYZ(data.frame(x = Switzerland$x,
								   y = Switzerland$y,
								   z = Switzerland$abundance))
tr <- transition(raster3, transitionFunction = f, directions = 8)
slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
points(high_abundance$x, high_abundance$y, pch = 16, cex = 1, col = adjustcolor("orange", 0.1)) 
for(j in 1:nrow(centroids)){
	for(k in 1:nrow(centroids)){
		plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
		     col = adjustcolor("cyan", 1), lwd = 1, add = TRUE)
	}
}
text(pts[,1], pts[,2], label = LETTERS[1:nrow(pts)], col = "green")

#align corridors with abundance estimates (posterior MCMC)
plot(raster, col = mapPalette(100), main = "Corridors with wildlife information (post. mcmc)")
points(high_abundance$x, high_abundance$y, pch = 16, cex = 1, col = adjustcolor("orange", 0.1)) 
for(i in 1:n){
	#stitch the pixels into a raster
	raster4 <- rasterFromXYZ(data.frame(x = Switzerland$x, y = Switzerland$y, z = tmp[i,]))
	#calculate paths connecting all centroids  
	tr <- transition(raster4, transitionFunction = f, directions = 8)
	slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
	for(j in 1:nrow(centroids)){
		for(k in 1:nrow(centroids)){
			plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
			     col = adjustcolor("cyan", 0.01), lwd = 1, add = TRUE)
		}
	}
}
text(pts[,1], pts[,2], label = LETTERS[1:nrow(pts)], col = "green")
```
\

## Notes

- swath of lines

When selecting proper alignment, corridor designers should examine a
series of lines and evaluate the advantages and disadvantages of moving to
the following cells rather than solely relying on the least-cost alignment.
This is because a least-cost corridor could be surrounded by otherwise
inappropriate habitats (e.g. human settlements) that are unlikely to be used and would be biologically
irrelevant. By incorporating the Bayesian method into this Least-Cost-Path
analysis, it gives not only the least-cost alignment (shown in magenta) but
other alignments (shown in cyan) as well. The latter can serve as alternatives
should any parts of the least-cost alignment cannot be used -- the more
progressively vivid cyan indicating the more preferred alignment to choose
from. 

```{r difference, echo = FALSE, warning = FALSE, message = FALSE}
# Combine post. mean and mcmc
plot(raster, col = mapPalette(100), main = "Least-cost corridors and alternatives")
points(high_abundance$x, high_abundance$y, pch = 16, cex = 1, col = adjustcolor("orange", 0.1)) 
for(i in 1:n){
	#stitch the pixels into a raster
	raster4 <- rasterFromXYZ(data.frame(x = Switzerland$x, y = Switzerland$y, z = tmp[i,]))
	#calculate paths connecting all centroids  
	tr <- transition(raster4, transitionFunction = f, directions = 8)
	slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
	for(j in 1:nrow(centroids)){
		for(k in 1:nrow(centroids)){
			plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
				 col = adjustcolor("cyan", 0.01), lwd = 1, add = TRUE)
		}
	}
}
raster3 <- rasterFromXYZ(data.frame(x = Switzerland$x,
								   y = Switzerland$y,
								   z = Switzerland$abundance))
tr <- transition(raster3, transitionFunction = f, directions = 8)
slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
for(j in 1:nrow(centroids)){
	for(k in 1:nrow(centroids)){
		plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
		     col = adjustcolor("magenta", 1), lwd = 2, add = TRUE)
	}
}
text(pts[,1], pts[,2], label = LETTERS[1:nrow(pts)], col = "green")
```

- anthropogenic resistance

Corridors often follow paths of least resistance based on the relationship
between species movement and features such as terrain, land cover, or
bioclimatic factors. Yet beyond the biophysical landscape, we also know that
human activity can influence species movement. What has not
considered in this analysis was the human activity context. Nevertheless,
framing it as "anthropogenic resistance" is also possible in the modeling, and
attributes such as human population densities and proximity from human
infrastructures could be added to the resistance surface.

```{r extras, eval = FALSE, include = FALSE, warning = FALSE, message = FALSE}
tmp2 <- matrix(NA, nrow = 5, ncol = nrow(Switzerland))
system.time(
for(i in 1:5){
	for(j in 1:nrow(Switzerland)){
		tmp2[i,j] <- exp(sim1[[3+i]][[1]] +
						 sim1[[3+i]][[2]]*Switzerland$forest[j] +
						 sim1[[3+i]][[3]]*Switzerland$forest[j]^2 +
						 sim1[[3+i]][[4]]*Switzerland$elevation[j] +
						 sim1[[3+i]][[5]]*Switzerland$elevation[j]^2)
	}
}
) # 1 sec

plot(raster, col = mapPalette(100), main = "Important corridors and alternatives")
for(i in 1:5){
	raster5 <- rasterFromXYZ(data.frame(x = Switzerland$x,
									   y = Switzerland$y,
									   z = tmp2[i,]))
	tr <- transition(raster5, transitionFunction = f, directions = 8)
	slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
	for(j in 1:nrow(centroids)){
		for(k in 1:nrow(centroids)){
			plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
				 col = adjustcolor("cyan", 1), lwd = 0.2, add = TRUE)
		}
	}
}
raster3 <- rasterFromXYZ(data.frame(x = Switzerland$x,
								   y = Switzerland$y,
								   z = Switzerland$abundance))
tr <- transition(raster3, transitionFunction = f, directions = 8)
slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
for(j in 1:nrow(centroids)){
	for(k in 1:nrow(centroids)){
		plot(shortestPath(slope, pts[j,], pts[k,], output = "SpatialLines"),
		     col = adjustcolor("magenta", 1), lwd = 2, add = TRUE)
	}
}
text(pts[,1], pts[,2], label = LETTERS[1:nrow(pts)], col = "green")
```


ddd <- list()
hhh <- list()
for(i in 1:1000){
	raster6 <- rasterFromXYZ(data.frame(x = Switzerland$x,
										y = Switzerland$y,
										z = tmp[i,]))
	tr <- transition(raster6, transitionFunction = f, directions = 8)
	slope <- geoCorrection(tr, type = "c", multpl = FALSE, scl = FALSE)
	sss <- shortestPath(slope, c(920000, 60000), c(1250000, 140000), output = "SpatialLines")
	ssss <- geom(sss)
	ddd[[i]] <- ssss[,4:5]
	fff <- pointDistance(ddd[[i]], lonlat = FALSE)
	ggg <- numeric(nrow(fff)-1)
	for(j in 1:(nrow(fff)-1)){
		ggg[j] <- fff[1+j, j]
	}
	hhh[[i]] <- ggg
}
jjj <- sapply(hhh, sum)
kkk <- max(jjj) - jjj
plot(raster, col = mapPalette(100), main = "Important corridors and alternatives")
colfunc <- colorRampPalette(c("white", "yellow", "orange", "red"))
ranking <- rank(kkk, ties.method = "first")
for(i in 1:length(kkk)){
	idx <- which(ranking == i)
	lines(ddd[[idx]], col = adjustcolor("red", 0.01)) #, col = colfunc(length(kkk))[i])
}
lines(ddd[[which.max(kkk)]], col = "green", lwd = 1)
points(c(920000, 1250000), c(60000, 140000), pch = 16, cex = 1, col = "green")
